### 2023-02-02 data_processing_LML.qmd
Initial data load

---
title: "Data cleaning and processing code"
author: "Leah Lariscy"
format: html
editor: visual
---

## Load necessary packages here

```{r}
library(tidyverse)
library(skimr)
```

## Load the data here

```{r}
# Load CFX N1 data
n1 <- read_csv("data/raw_data/n1_data.csv")
# Load CFX N2 data
n2 <- read_csv("data/raw_data/n2_data.csv")
# Load Plant data
plant <- read_csv("data/raw_data/plant_data.csv")
# Load Standard curve data
std_curve <- read_csv("data/raw_data/standard_curves.csv")
# Load 
read_csv("data/raw_data/")
```