### 2023-02-03 code/processing_code/data_processing_LML.qmd
part 1 ready

---
title: "Data cleaning and processing script"
author: "Leah Lariscy"
format: html
editor: visual
---

## Data introduction

For this project, I would like to work with my own real-world data to explore the relationship between SARS-CoV-2 waste water surveillance data and COVID-19 clinical case data.

### Hypotheses so far

Waste water surveillance data will begin to correlate less with the clinical data as we progress further into the pandemic. (as rapid, at-home testing becomes more prevalent)

Catchment-level surveillance data will correlate better to waste water catchment-level case data than it will with county-level data.

## Load necessary packages here

```{r}
library(tidyverse)
library(skimr)
library(here)
```

## Load the data here

```{r}
here()
# Load CFX N1 data
n1 <- read_csv(here("data/raw_data/cfx_n1.csv"))

# Load CFX N2 data
n2 <- read_csv(here("data/raw_data/cfx_n2.csv"))

# Load Plant data
plant <- read_csv(here("data/raw_data/plant_data.csv"))

# Load COVID-19 Symptom data
covid_symptom <- read_csv(here("data/raw_data/ga_covid_data/epicurve_symptom_date.csv")) %>% 
  filter(county=="Clarke") %>% 
  select(symptom.date=`symptom date`, 
         cases, moving_avg_cases)

#Load COVID-19 Confirmed Case Data
covid_case <- read_csv(here("data/raw_data/ga_covid_data/epicurve_rpt_date.csv")) %>% 
  filter(county=="Clarke") %>% 
  select(report_date, 
         cases, 
         moving_avg_cases)

#Load COVID-19 Testing Data
covid_testing <- read_csv(here("data/raw_data/ga_covid_data/pcr_antigen_col.csv")) %>% 
  filter(county=="Clarke") %>% 
  select(collection_date = collection_dt, 
         pcr_tests = `ALL PCR tests performed`, 
         pcr_pos = `All PCR positive tests`, 
         pcr_pos_7dma = `7 day percent positive`,
         pcr_pos_14dma = `14 day percent positive`)

#Load Catchment-level COVID-19 Case Data
#covid_case_wrf = read_csv(here("data/raw_data/cases_wrf.csv"))
```

## Description of the data

### CFX N1 Data

`n1` \<- Waste water extracted RNA RT-qPCR results of SARS-CoV-2 N1 gene target

```{r}
skim(n1)
```

### CFX N2 Data

`n2` \<- Waste water extracted RNA RT-qPCR results of SARS-CoV-2 N2 gene target

```{r}
skim(n2)
```

### Plant Data

`plant` \<- Flow data from three Water Reclamation Facilities (WRF) in Athens-Clarke County (ACC).

```{r}
skim(plant)
```

### County-level COVID-19 Symptom Data

`covid_symptom` \<- Count of individuals who reported having symptoms of COVID-19 on a given day, subset to include only data from ACC. Accessed from Georgia Department of Public Health (GDPH) on 2023-01-31.

```{r}
skim(covid_symptom)
```

### County-level COVID-19 Case Data

`covid_case` \<- Count of confirmed cases of COVID-19 on a given day, subset to include only data from ACC. Accessed from GDPH on 2023-01-31.

```{r}
skim(covid_case)
```

### County-level COVID-19 Testing Data

`covid_testing` \<- Count of COVID-19 tests administered on a given day, subset to include only data from ACC. Accessed from GDPH on 2023-01-31.

```{r}
skim(covid_testing)
```

### Catchment-level COVID-19 Case Data

`covid_case_wrf` \<- Count of confirmed cases of COVID-19 on a given day from each WRF catchment zone in ACC.

```{r}
#I don't have access to this data yet, waiting on DPH partner to send.
```