### 2023-02-23 code/analysis_code/exploratory_analysis_LML.qmd
exploratory analysis

---
title: "Exploratory analysis of year 2 surveillance data"
author: "Leah Lariscy"
format: html
editor: visual
theme: solar
---

## Load necessary packages

```{r}
library(tidyverse)
library(here)
library(ggpmisc)

```

## Load data frames generated in data processing script

```{r}
here()
n1_n2_clean <- readRDS(here("data/processed_data/year2_cfx_clean.rds"))
plant_clean <- readRDS(here("data/processed_data/year2_plant_clean.rds"))
n1_n2_plant_combine <- readRDS(here("data/processed_data/year2_cfx_plant_combined.rds"))
covid_combine <- readRDS(here("data/processed_data/all_covid_combined.rds"))
all_data_combine <- readRDS(here("data/processed_data/year2_all_data_combined.rds"))
recovery_percent <- readRDS(here("data/processed_data/recovery_percentages.rds"))

```

### Look at `all_data_combine` and determine visualization to make

```{r}
tibble(all_data_combine)
```

## Data Viz

Some plots I want to generate initially are

1.  Case averages over time
2.  Total copies over time (N1 and N2)

### Look at case average over time

```{r}
all_data_combine %>% group_by(date) %>% 
  summarize(mean_7dma = mean(pcr_pos_7dma)) %>% 
  ggplot(aes(x=date, y=mean_7dma)) +
  geom_point() +
 stat_smooth(method = "gam")

```

### Look at log10 total copies in wastewater over time for N1

First, average total copies of biological replicates. Then, add total copies from each facility.

```{r}
all_data_combine %>% filter(target == "N1") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_total_copies = mean(total_copies)) %>% 
  ungroup(facility) %>% 
  summarize(sum_total_copies = sum(mean_total_copies)) %>% 
  ggplot(aes(x = date,y = log10(sum_total_copies))) +
  geom_point(color = "red") +
  stat_smooth(method = "gam")

```

### Look at log10 total copies in wastewater over time for N2

```{r}
all_data_combine %>% filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_total_copies = mean(total_copies)) %>% 
  ungroup(facility) %>% 
  summarize(sum_total_copies = sum(mean_total_copies)) %>% 
  ggplot(aes(date, log10(sum_total_copies))) +
  geom_point() +
  stat_smooth(method = "gam")
```

### Log10 total copies per day for N1 and N2

```{r}
all_data_combine %>% 
  drop_na() %>% 
  group_by(date, target, facility) %>% 
  summarize(mean_total_copies = mean(total_copies)) %>% 
  ungroup(facility) %>% 
  summarize(sum_total_copies = sum(mean_total_copies)) %>% 
  ggplot(aes(date, log10(sum_total_copies), color = target)) +
  geom_point() +
  stat_smooth(method = "gam")
```

### Look at log10 copies per L in N1

```{r}
all_data_combine %>% filter(target == "N1") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(copy_num_L)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L)) %>% 
  ggplot(aes(date, log10(sum_mean_copies_L))) +
  geom_point() +
  stat_smooth(method = "gam")
```

### Look at log10 copies per liter in N1

```{r}
all_data_combine %>% filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(copy_num_L)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L)) %>% 
  ggplot(aes(date, log10(sum_mean_copies_L))) +
  geom_point() +
  stat_smooth(method = "gam")
```

### Look at log10 copies per liter across both N1 and N2

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(copy_num_L)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L)) %>% 
  ggplot(aes(date, log10(sum_mean_copies_L))) +
 stat_smooth(method = "gam")
```

### Look at log10 copies per liter vs case 7dma

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(copy_num_L), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(log10(sum_mean_copies_L), mean_7dma)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq()
```

### Look at log10 copies of N1 per liter vs case 7dma

```{r}
all_data_combine %>% filter(target == "N1") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(copy_num_L), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(log10(sum_mean_copies_L), mean_7dma)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq()
```

### Look at log10 copies of N2 per liter vs case 7dma

```{r}
all_data_combine %>% filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(copy_num_L), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(log10(sum_mean_copies_L), mean_7dma)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq()
```

### Look at relationship of copies/L and 7dma over time

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(copy_num_L), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(x=date)) +
  #geom_point(aes(y=log10(sum_mean_copies_L)/mean_7dma)) + 
  geom_line(aes(y=(log10(sum_mean_copies_L)/mean_7dma), color = "red")) +
  geom_line(aes(y=log10(mean_7dma)), color = "blue")

```