### 2023-03-13 products/slides/Slides.qmd

---
title: Example Quarto slides
author: Andreas Handel
format: revealjs
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

## Overview

-   A few simple slides using the [`revealjs` (html) format](https://quarto.org/docs/presentations/revealjs/).
-   For other formats (e.g. Powerpoint, or Beamer/pdf), see [here](https://quarto.org/docs/guide/).

## Example slide

This shows the summary table. It is pulled in as R chunk.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
#resulttable=readRDS(here::here("results","summarytable.rds"))
#knitr::kable(resulttable)
```

## Example slide

This shows a figure created by the analysis script. It is inserted using Markdown (not knitr code, but that would be possible too).

![](images/height_weight.png){fig-align="center" width="420"}

## Example slide

This shows the model fitting results as table.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
#resulttable2 = readRDS(here::here("results","resulttable2.rds"))
#knitr::kable(resulttable2)
```

## Example slide with reference

This paper [@leek2015] discusses types of analyses.

## References