### 2023-04-24 code/analysis_code/exploratory/exploratory_analysis.qmd
fixing issues

---
title: "Exploratory analysis of year 2 surveillance data"
author: "Leah Lariscy"
editor: visual
format:
  html:
    toc: true
    toc-depth: 3
theme: solar
---

## Load necessary packages

```{r, message=FALSE}
library(tidyverse)
library(here)
library(ggpmisc)
library(ggthemes)

```

## Load final data frame generated in data processing script

```{r}
data_clean <- readRDS(here("data/processed_data/final_sum_combined_data.RDS"))

```

### Look at `data_clean`

```{r}
tibble(data_clean)
```

# Data Viz

### Variation analysis

-   histogram of case count observations

-   histogram of 7 day moving average case observations

-   histogram of total daily copies

-   histogram of copies per L

-   histogram of percent positive qPCR assays

### Co-variation analysis

-   Scatter plot of wastewater and clinical variables over time

-   Linear regressions between wastewater and clinical variables

## Variation analysis

### Clinical case 7 day moving average histogram

```{r}
data_clean %>% ggplot(aes((case_7dma))) +
  geom_histogram(fill = "#C06C84") +
xlab(label = "Clinical Case Seven Day Moving Average") +
ylab(label = "Count") +
theme_light()

ggsave(here("figures/exploratory/variate/7DMA_histo.png"))
summary(data_clean$case_7dma)
```

Slight bimodal distribution in 7DMA

### PCR positive clinical cases histogram

```{r}
data_clean %>% ggplot(aes(pcr_pos)) +
  geom_histogram(fill = '#F8B195') +
  xlab(label = "Daily Positive Clinical Cases") +
  ylab(label = "Count") +
  theme_light()

ggsave(here("figures/exploratory/variate/pos_cases_histo.png"))
summary(data_clean$pcr_pos)
```

The distribution of daily cases is heavily skewed to the right. Log-transforming gives a more normal distribution

### Clinical case percent positive tests

```{r}
data_clean %>% ggplot(aes(percent_pos_test)) +
  geom_histogram(fill = "#FF7E5A")

summary(data_clean$percent_pos_test)

ggsave(here("figures/exploratory/variate/14DMA_histo.png"))

```

There is also a slight bimodal distribution with percent of positive tests, but nonetheless the distribution is right-skewed.

### Wastewater viral load histogram

```{r}
    data_clean %>% ggplot(aes(avg_VL)) +
  geom_histogram() +
    xlab(label = "Wastewater viral load") +
    ylab(label = "Count") +
    theme_light()

ggsave(here("figures/exploratory/variate/VL_histo.png"))

```

Wastewater viral load is also very right-skewed. We may need to log-transform some variables before doing more statistical analysis.

### Log10 viral load histogram

```{r}
data_clean %>% 
  ggplot(aes(log10_VL)) +
  geom_histogram()

ggsave(here("figures/exploratory/variate/log10VL_histo.png"))

```

### Assay percent positivity histogram

```{r}
  data_clean %>% ggplot(aes(pos_rate_tr)) +
  geom_histogram() +
    xlab(label = "qPCR assay positivity") +
    ylab(label = "Count") +
    theme_light()

ggsave(here("figures/exploratory/variate/assay_pos_histo.png"))

```

Again with the bimodal distribution, even more so this time.

## Co-variate analysis: time series

### Case 7 day moving average overtime

```{r, message=FALSE}
data_clean %>% 
  ggplot(aes(x = date, y = case_7dma)) +
  geom_line(size =1.5, color = '#4c005c') +
  xlab(label = "Date") +
  ylab(label = "Clinical Case 7-Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/7DMA_time_plot.png"))

```

### PCR positives per day over time

```{r}
data_clean %>% 
  ggplot(aes(x = date, y = pcr_pos)) +
  geom_line(color = '#4c005c', size = 2) +
  ylab(label = "Positive Cases Per Day") +
  xlab(label = "Date") +
  theme_light()

ggsave(here("figures/exploratory/covariate/pos_cases_time.png"))
```

### Clinical case percent positive tests over time

```{r}
data_clean %>% 
   ggplot(aes(x = date, y = percent_pos_test)) +
  geom_line(size = 1.5) +
  ylab(label = "Percent of positive clinical tests") +
  xlab(label = "Date") +
  theme_light()

ggsave(here("figures/exploratory/covariate/14DMA_time.png"))
```

### Wastewater viral load over time

```{r}
data_clean %>% 
  ggplot(aes(date, avg_VL)) +
  geom_line(color = "#2bce48", size = 1.5) +
   xlab(label = "Date") +
  ylab(label = "Total Daily County-Level Copies") +
  theme_light()

ggsave(here("figures/exploratory/covariate/avg_total_copies_time.png"))
```

### Log10 wastewater viral load over time

```{r}
data_clean %>% 
  ggplot(aes(date, log10_VL)) +
  geom_line(color = "#2bce48", size = 1.5) +
   xlab(label = "Date") +
  ylab(label = "Log10 Total Daily County-Level Copies") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_time.png"))
```

### Percent positive technical replicates per sample over time

```{r, message=FALSE}
data_clean %>% 
  ggplot(aes(date, pos_rate_tr)) +
geom_line(size =1.5, color = '#634499') +
  xlab(label = "Date") +
  ylab(label = "Percent Positive N1 & N2 assays") +
  theme_light()

ggsave(here("figures/exploratory/covariate/percent_pos_time_plot.png"))

```

### Tests administered over time

```{r}
data_clean %>% ggplot(aes(date, test_admin)) +
  geom_line(size = 1.5) +
  theme_light()

ggsave(here("figures/exploratory/covariate/test_admin_time.png"))
```

Tests administered drop off significantly after the beginning of 2022. In all other time series, there is a spike in mid-2022, but that spike is missing here.

From here on, I will split the data before and after Februrary 28, 2022 to see how correlations change.

## Co-variate analysis: correlations between WW and clinical variables

### Log10 wastewater viral load vs case 7DMA

```{r}
#full dataset
data_clean %>% 
  ggplot(aes(log10_VL, case_7dma)) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
   xlab(label = "Log10 County-Level Copies Per Liter") +
  ylab(label = "Log10 Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_vs_7DMA.png"))

#data split before FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>% 
  ggplot(aes(log10_VL, case_7dma)) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
   xlab(label = "Log10 County-Level Copies Per Liter") +
  ylab(label = "Log10 Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_vs_7DMA_Y1.png"))

#data split after FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>% 
  ggplot(aes(log10_VL, case_7dma)) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
   xlab(label = "Log10 County-Level Copies Per Liter") +
  ylab(label = "Log10 Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_vs_7DMA_Y2.png"))
```

RSQ value does not change between full data and the subset of dates

### Log10 wastewater viral load vs tests administered

```{r}
#full dataset
data_clean %>%  
  ggplot(aes(log10_VL, test_admin)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "Log10 viral load") +
  ylab(label = "PCR tests administered") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_testadmin.png"))

#data split before FEB 28 2022
data_clean %>%  
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>% 
  ggplot(aes(log10_VL, test_admin)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "Log10 viral load") +
  ylab(label = "PCR tests administered") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_testadmin_Y1.png"))

#data split after FEB 28 2022
data_clean %>%  
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>% 
  ggplot(aes(log10_VL, test_admin)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "Log10 viral load") +
  ylab(label = "PCR tests administered") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_testadmin_Y2.png"))
```

Correlations overall are weak, but the strongest positive correlation is found in the earlier dates

### Log10 wastewater viral load vs percent of positive tests

```{r}
#full dataset
data_clean %>%  
  ggplot(aes(log10_VL, percent_pos_test)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "Log10 wastewater viral load") +
  ylab(label = "Percent of positive clinical cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_percentcase.png"))

#data split before FEB 28 2022
data_clean %>%  
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>%
  ggplot(aes(log10_VL, percent_pos_test)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "Log10 wastewater viral load") +
  ylab(label = "Percent of positive clinical cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_percentcase_Y1.png"))

#data split after FEB 28 2022
data_clean %>%  
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>%
  ggplot(aes(log10_VL, percent_pos_test)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "Log10 wastewater viral load") +
  ylab(label = "Percent of positive clinical cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_percentcase_Y2.png"))
```

RSQ is slightly lower in later dates than in full data set, but does not change in earlier dates

### Log10 wastewater viral load vs daily reported cases

```{r}
#full dataset
data_clean %>%  
  ggplot(aes(log10_VL, log10(pcr_pos))) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
  xlab(label = "Log10 wastwater viral load") +
  ylab(label = "Daily Positive Clinical Cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_dailycases.png"))

#data split before FEB 28 2022
data_clean %>%  
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>%
  ggplot(aes(log10_VL, log10(pcr_pos))) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
  xlab(label = "Log10 wastwater viral load") +
  ylab(label = "Daily Positive Clinical Cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_dailycases_Y1.png"))

#data split after FEB 28 2022
data_clean %>%  
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>%
  ggplot(aes(log10_VL, log10(pcr_pos))) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
  xlab(label = "Log10 wastwater viral load") +
  ylab(label = "Daily Positive Clinical Cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_dailycases_Y2.png"))
```

RSQ is higher is the data splits than in the full set. I'm not sure what to think of this, but it could be due to the high error associated with this metric.

### Assay positivity vs 7DMA

```{r}
#full dataset
data_clean %>% 
  ggplot(aes(pos_rate_tr, case_7dma)) +
  geom_point(color = "#00b0ba") +
  stat_poly_line(color = "#00b0ba") +
  stat_poly_eq(color = "black") +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_vs_7DMA.png"))

#data split before FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>%
  ggplot(aes(pos_rate_tr, case_7dma)) +
  geom_point(color = "#00b0ba") +
  stat_poly_line(color = "#00b0ba") +
  stat_poly_eq(color = "black") +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_vs_7DMA_Y1.png"))

#data split after FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>%
  ggplot(aes(pos_rate_tr, case_7dma)) +
  geom_point(color = "#00b0ba") +
  stat_poly_line(color = "#00b0ba") +
  stat_poly_eq(color = "black") +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_vs_7DMA_Y2.png"))
```

No significant change in data splits

### Assay positivity vs tests administered

```{r}
#full dataset
data_clean %>% 
  ggplot(aes(pos_rate_tr,test_admin)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "PCR tests administered") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_testadmin.png"))

#data split before FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>%
  ggplot(aes(pos_rate_tr,test_admin)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "PCR tests administered") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_testadmin_Y1.png"))

#data split after FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>%
  ggplot(aes(pos_rate_tr,test_admin)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "PCR tests administered") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_testadmin_Y2.png"))
```

RSQ is low across all data sets, but highest in earlier dates than in full data or later dates

### Assay positivity vs percent of positive tests

```{r}
#full dataset
data_clean %>% 
  ggplot(aes(pos_rate_tr,percent_pos_test)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "PCR test positivity") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_testpos.png"))

#data split before FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>%
  ggplot(aes(pos_rate_tr,percent_pos_test)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "PCR test positivity") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_testpos_Y1.png"))

#data split after FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>%
  ggplot(aes(pos_rate_tr,percent_pos_test)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "PCR test positivity") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_testpos_Y2.png"))
```

There is a slight decrease in RSQ in the later dates than in the full set or the earlier dates.

### Assay positivity vs daily reported cases

```{r}
#full dataset
data_clean %>% 
  ggplot(aes(pos_rate_tr,pcr_pos)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "Daily reported cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_dailycases.png"))

#data split before FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>%
  ggplot(aes(pos_rate_tr,pcr_pos)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "Daily reported cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_dailycases_Y1.png"))

#data split after FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>%
  ggplot(aes(pos_rate_tr,pcr_pos)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "Daily reported cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_dailycases_Y2.png"))
```

The highest RSQ value was in the earlier dates than in the full set or later dates