### 2023-04-24 code/processing_code/percent_pos.Rmd

---
title: "Calculating percent positive assays"
author: "Leah Lariscy"
date: "2023-03-26"
output: html_document
---

## Load packages

```{r}
library(tidyverse)
library(here)
```

## Load data

```{r}
wbe <- readRDS(here("data/processed_data/year2_cfx_clean.rds"))
```

## Calculate percent positive qPCR assays

```{r}
wbe_pos <- wbe %>% group_by(date, target, facility, biological_replicate) %>% 
  #n positive per biological rep
  summarise(n=n(),
            n.miss=sum(is.na(ct)),
            n.pos=n-n.miss) %>% 
  #n positive per target/facility/date
  ungroup(biological_replicate) %>% 
  summarise(n=sum(n),
            n.miss=sum(n.miss),
            n.pos=n-n.miss) %>% 
  #n positive per target/date
  ungroup(facility) %>% 
  summarise(n=sum(n),
            n.miss=sum(n.miss),
            n.pos=n-n.miss) %>%
  ungroup()

wbe_pos
```

### Separate N1 and N2

```{r}
wbe_pos_n1 <- wbe_pos %>% filter(target == "N1")

wbe_pos_n2 <- wbe_pos %>% filter(target == "N2")

saveRDS(wbe_pos_n1, here("data/processed_data/percent_pos_n1.rds"))
saveRDS(wbe_pos_n2, here("data/processed_data/percent_pos_n2.rds"))
```

### Sum N1 and N2

```{r}
# n positive per date
wbe_pos_both <- wbe_pos %>% group_by(date) %>% 
   summarise(n=sum(n),
            n.miss=sum(n.miss),
            n.pos=n-n.miss) %>% 
  mutate(percent_pos = (n.pos/n)*100) 
wbe_pos_both
```

### Avg N1 and N2

```{r}
wbe_pos_avg <- wbe_pos %>% group_by(date) %>% 
   summarise(n=sum(n),
            n.miss=sum(n.miss),
            n.pos=n-n.miss) %>% 
  mutate(percent_pos = (n.pos/n)*100) 
wbe_pos_avg
```

## Load COVID data and combine with qPCR data

```{r}
covid <- readRDS(here("data/processed_data/all_covid_combined.rds"))

covid_n1 <- full_join(wbe_pos_n1, covid, by = "date") %>% drop_na(target)

covid_n2 <- full_join(wbe_pos_n2, covid, by = "date") %>% drop_na(target)

covid_both <- full_join(wbe_pos_both, covid, by = "date") %>% drop_na(n)

saveRDS(covid_n1, here("data/processed_data/covid_ww_n1.rds"))
saveRDS(covid_n2, here("data/processed_data/covid_ww_n2.rds"))
saveRDS(covid_both, here("data/processed_data/covid_ww_both.rds"))
```

```{r}
covid_both %>% ggplot(aes(date, percent_pos)) +
geom_line(size =1.5, color = '#634499') +
  xlab(label = "Date") +
  ylab(label = "Percent Positive qPCR assays") +
  theme_light()

ggsave(here("figures/exploratory/covariate/percent_pos_time_plot.png"))
```

```{r}
  covid_both %>% 
    filter(between(date, as.Date("2020-06-30"), as.Date("2022-12-21"))) %>% 
    ggplot(aes(percent_pos, pcr_pos_7dma)) +
  geom_point(color = "#634499") +
  stat_poly_line(color = "#634499") +
  stat_poly_eq(color = "black") +
  xlab(label = "Percent Positve N1 & N2 Assays") +
  ylab(label = "Clinical Case Seven Day Moving Average") +
  theme_light()

  ggsave(here("figures/exploratory/covariate/PP_vs_7DMA.png"))
```

```{r}
 covid_both %>% 
    filter(between(date, as.Date("2020-06-30"), as.Date("2021-06-30"))) %>% 
    ggplot(aes(percent_pos, pcr_pos)) +
  geom_point(color = "#4dd091") +
  stat_poly_line(color = "#4dd091") +
  stat_poly_eq(color = "black") +
  xlab(label = "Percent Positve N1 & N2 Assays") +
  ylab(label = "Daily Reported Clinical Cases") +
  theme_light()

 ggsave(here("figures/exploratory/covariate/PP_daily_cases.png"))
```