### 2023-04-24 code/processing_code/processing.qmd

---
title: "Data cleaning and processing script"
author: "Leah Lariscy"
editor: visual
format:
  html:
    toc: true
    toc-depth: 3
theme: solar
---

## Data introduction

For this project, I would like to work with my own real-world data to explore the relationship between **SARS-CoV-2 waste water surveillance data** and **COVID-19 clinical case data**.

## Questions I have so far

### Variation

1.  What is the distribution like of SARS-CoV-2 viral load observations? What does this distribution look like at catchment levels vs at the county level?
2.  What is the distribution like of N1 and N2 qPCR percent positive observations? What does this distribution look like at catchment levels vs at the county level?
3.  What is the distribution like of clinical case observations? What does this distribution look like at catchment levels vs at the county level?

### Covariation

1.  Is there a relationship between viral load and clinical case observations?

2.  Is there a relationship between qPCR percent positive and clinical case observations?

3.  What do viral load trends look like over time?

4.  What do clinical case trends look like over time?

5.  How do each of these temporal trends compare to one another over time?

## Load necessary packages here

```{r}
library(tidyverse)
library(skimr)
library(here)
```

## Load the data here

```{r, message=FALSE}
here()
# Load CFX N1 data
n1_1 <- read_csv(here("data/raw_data/n1_data_year1.csv")) #year 1 data
n1 <- read_csv(here("data/raw_data/cfx_n1.csv")) #year 2 data

# Load CFX N2 data
n2_1 <- read_csv(here("data/raw_data/n2_data_year1.csv")) #year 1 data
n2 <- read_csv(here("data/raw_data/cfx_n2.csv")) #year 2 data

# Load Plant data
plant <- read_csv(here("data/raw_data/plant_data.csv"))

# Load COVID-19 Symptom data
covid_symptom <- read_csv(here("data/raw_data/ga_covid_data/epicurve_symptom_date.csv")) %>% 
  filter(county=="Clarke") %>% 
  select(symptom.date=`symptom date`, 
         cases, moving_avg_cases)

#Load COVID-19 Confirmed Case Data
covid_case <- read_csv(here("data/raw_data/ga_covid_data/epicurve_rpt_date.csv")) %>% 
  filter(county=="Clarke") %>% 
  select(report_date, 
         cases, 
         moving_avg_cases)

#Load COVID-19 Testing Data
covid_testing <- read_csv(here("data/raw_data/ga_covid_data/pcr_antigen_col.csv")) %>% 
  filter(county=="Clarke") %>% 
  select(collection_date = collection_dt, 
         pcr_tests = `ALL PCR tests performed`, 
         pcr_pos = `All PCR positive tests`, 
         pcr_pos_7dma = `7 day percent positive`,
         pcr_pos_14dma = `14 day percent positive`)

#Load Catchment-level COVID-19 Case Data
#covid_case_wrf = read_csv(here("data/raw_data/cases_wrf.csv"))

#Load CFX recovery data
recovery_output <- read_csv(here("data/raw_data/recovery_data.csv"))
recovery_input <- read_csv(here("data/raw_data/calfguard.csv"))
```

## Description of the data

### CFX N1 Data

`n1` \<- Waste water extracted RNA RT-qPCR results of SARS-CoV-2 N1 gene target

```{r}
skim(n1)
```

### CFX N2 Data

`n2` \<- Waste water extracted RNA RT-qPCR results of SARS-CoV-2 N2 gene target

```{r}
skim(n2)
```

### Plant Data

`plant` \<- Flow data from three Water Reclamation Facilities (WRF) in Athens-Clarke County (ACC).

```{r}
skim(plant)
```

### County-level COVID-19 Symptom Data

`covid_symptom` \<- Count of individuals who reported having symptoms of COVID-19 on a given day, subset to include only data from ACC. Accessed from Georgia Department of Public Health (GDPH) on 2023-01-31.

```{r}
skim(covid_symptom)
```

### County-level COVID-19 Case Data

`covid_case` \<- Count of confirmed cases of COVID-19 on a given day, subset to include only data from ACC. Accessed from GDPH on 2023-01-31.

```{r}
skim(covid_case)
```

### County-level COVID-19 Testing Data

`covid_testing` \<- Count of COVID-19 tests administered on a given day, subset to include only data from ACC. Accessed from GDPH on 2023-01-31.

```{r}
skim(covid_testing)
```

## Data cleaning and wrangling

Currently, I have 6 datasets. In order to do analysis on these data down the road, I want to figure out how to best combine them.

Here is a list of the data sets I have:

1.  CFX N1 results
2.  CFX N2 results
3.  WWTP flow rate and total solids
4.  County-level COVID symptoms
5.  County-level COVID confirmed cases
6.  County-level COVID tests administered

I will start by combining the CFX output data with the plant flow data and calling the data frame `wbe`

Next, I will combine all COVID-related data

### Remove unnecessary variables

```{r}
#N1
n1_1 <- n1_1 %>% select(c(sample_date, collection_num, sample_id, target, ct))
n1 <- n1 %>% select(c(sample_date, collection_num, sample_id, target, ct)) %>% 
  filter(between(sample_date, as.Date("2021-07-06") , as.Date("2022-12-21")))

#N2
n2_1 <- n2_1 %>% select(c(sample_date, collection_num, sample_id, target, ct))
n2 <- n2 %>% select(c(sample_date, collection_num, sample_id, target, ct)) %>% 
  filter(between(sample_date, as.Date("2021-07-06") , as.Date("2022-12-21")))
```

### Calculate SARS-CoV-2 concentrations from CFX data for both N1 and N2

n1 \<- n1 %\>% mutate(ct = as.numeric(ct),

Year1:

Year 2: copy_num_uL_rxn = as.numeric(10\^((ct-36.046)/-3.5293)), #use standard curve slope here

copy_num_uL_rxn = replace_na(copy_num_uL_rxn, 0.004), #replace NAs with limit of detection (0.004 copies per uL of the reaction)

run_date = as.Date(run_date))

```{r}
#year 1 N1 data here
n1_1 <- n1_1 %>% mutate(ct = as.numeric(ct), 
                copy_num_uL_rxn = as.numeric(10^((ct-34.008)/-3.389)),
                copy_num_uL_rxn = replace_na(copy_num_uL_rxn, 0.004))

#year 2 N1 data here
n1 <- n1 %>% mutate(ct = as.numeric(ct), 
              copy_num_uL_rxn = as.numeric(10^((ct-36.046)/-3.5293)), #use standard curve slope here
              copy_num_uL_rxn = replace_na(copy_num_uL_rxn, 0.004)) #replace NAs with limit of detection (0.004 copies per uL of the reaction)

#N2 year 1 data here
n2_1 <-n2_1 %>% mutate(ct = as.numeric(ct),
                copy_num_uL_rxn = as.numeric(10^((ct-32.416)/-3.3084)),
                copy_num_uL_rxn = replace_na(copy_num_uL_rxn, 0.004))

#N2 year 2 data here
n2 <- n2 %>% mutate(ct = as.numeric(ct),
              copy_num_uL_rxn = as.numeric(10^((ct-37.731)/-3.2505)),
              copy_num_uL_rxn = replace_na(copy_num_uL_rxn, 0.004))

```

### Combining year 1 and 2 data

```{r}
#N1
n1 <- bind_rows(n1_1, n1)
tibble(n1)

#N2
n2 <- bind_rows(n2_1, n2)
tibble(n2)
```

### Combining and cleaning CFX data

`wbe_year2` includes: date, facility, target, bio rep, collection number, ct, copies/uL/rxn, copies/L

```{r}

wbe_year2 <- bind_rows(n1, n2) %>% #combine dataframes with same variables
  mutate(
    date=as.Date(sample_date, format = "%d-%b-%y"), #change to date format
    facility=substr(sample_id, 1,2), #first two letters in sample_id is facility ID
    biological_replicate=substr(sample_id, nchar(sample_id), nchar(sample_id)), #last number in sample_id is the replicate number
    ct=as.numeric(ifelse(ct=="Undetermined", NA, ct)) #set NAs
  ) %>% 
  arrange(date, facility, target, biological_replicate) %>% 
  select(date, facility, target, biological_replicate, collection_num, ct, copy_num_uL_rxn) %>%
  mutate(facility = as.factor(facility), 
         facility = recode(facility, NO = "A", MI = "B", CC = "C"), 
         facility = ordered(facility, levels = c("A", "B", "C")), #set facility to factor and then set factor levels
         copy_num_L = copy_num_uL_rxn*20/5*60/280*1000*1000)

saveRDS(wbe_year2, here("data/processed_data/year2_cfx_clean.rds"))
wbe_year2
```

### Summarizing CFX data for downstream processing

`wbe_year2_avg` - copies/L is averaged across biological replicates

```{r}
wbe_year2_avg <- wbe_year2 %>% group_by(date,facility,target,collection_num) %>% 
  summarize(avg_copy_L = mean(copy_num_L), 
            avg_copy_uL_rxn = mean(copy_num_uL_rxn)) %>% 
  ungroup()
saveRDS(wbe_year2_avg, here("data/processed_data/year2_cfx_avg.rds"))
wbe_year2_avg
```

### Cleaning plant data

```{r}
plant_clean <- plant %>% 
  mutate(date = as.Date(date), #change to date format
         influent_flow_L = influent_flow_mg*1e6*231*(0.0254^3)*1000) %>%  #convert flow from millions of gallons to liters
  mutate(facility = as.factor(wrf), facility = recode(facility, NO = "A", MI = "B", CC = "C"), facility = ordered(facility, levels = c("A", "B", "C"))) %>% #set facility to factor and then set factor levels
   select(date, facility, influent_flow_L, influent_tss_mg_l)

saveRDS(plant_clean, here("data/processed_data/year2_plant_clean.rds"))
plant_clean
```

### Combining CFX and plant data

```{r}
n1_n2_plant <- left_join(wbe_year2, plant_clean, by = c("date", "facility")) %>% 
  mutate(total_copies = copy_num_L * influent_flow_L)

saveRDS(n1_n2_plant, here("data/processed_data/year2_cfx_plant_combined.rds"))
n1_n2_plant
```

### Combining all COVID data

```{r}
covid <- full_join(
  covid_symptom%>%
    select(cases.symptom.onset=cases, date=symptom.date), 
  covid_case%>%
    select(cases.reported=cases, date=report_date), 
  by = "date"
) %>% 
  full_join(
    covid_testing%>%
      rename(date=collection_date), 
    by="date"
  ) %>%
  select(date, cases.symptom.onset, cases.reported, pcr_tests, pcr_pos, pcr_pos_7dma, pcr_pos_14dma) %>%

  mutate(prop_pos = pcr_pos/pcr_tests)

saveRDS(covid, here("data/processed_data/all_covid_combined.rds"))
covid
```

### Combining `n1_n2_plant` and `covid`

```{r}
all_data <- full_join(n1_n2_plant, covid, by = "date") %>% 
   filter(between(date, as.Date("2020-06-30") , as.Date("2022-12-21")))

saveRDS(all_data, here("data/processed_data/year2_all_data_combined.rds"))
all_data
```

### Cleaning recovery data sets

```{r}
#Start with recovery_output 
recovery_output_clean <- recovery_output %>% 
  mutate(ct = as.numeric(ct)) %>% 
  group_by(sample_id, cg_num) %>% 
  summarize(avg_ct = mean(ct)) %>% 
  separate(sample_id, into = c("facility", "collection_num", "rep"), sep = "_") %>% 
  mutate(collection_num = as.numeric(collection_num),
         copies_ul_rxn = 10^((avg_ct-30.7)/-3.238),
         copies_ul_sample = copies_ul_rxn *20/2*25/3*60/280) %>% 
  select(facility, collection_num, cg_num, copies_ul_sample, avg_ct) %>% 
  drop_na()

recovery_input_clean <- recovery_input %>% 
  mutate(ct = as.numeric(ct)) %>% 
  group_by(sample_id) %>% 
  summarize(avg_ct = mean(ct)) %>% 
  separate(sample_id, into = c("CG", "cg_num"), sep = "_") %>% 
  mutate(cg_num = as.numeric(cg_num),
         copies_ul_rxn = 10^((avg_ct-30.7)/-3.238),
         copies_ul_input = copies_ul_rxn *20/2*25/3*60/50*40/40000) %>% 
  select(cg_num, copies_ul_input, avg_ct) %>% 
  drop_na()

saveRDS(recovery_output_clean, here("data/processed_data/recovery_output_clean.rds"))
saveRDS(recovery_input_clean, here("data/processed_data/recovery_input_clean.rds"))

```

### Calculate percent positive of technical reps

This isn't super important, but I'm including anyways.

This will calculate the percent of positive technical replicates per biological replicate

```{r}
wbe_tr <- wbe_year2 %>% group_by(date, facility, target, biological_replicate) %>% 
  dplyr::summarise(
    n=n(), 
    n.miss=sum(is.na(ct)), 
    ct.mean=mean(ct,na.rm=T), 
    ct.sd=sd(ct,na.rm=T)
  ) %>% 
  mutate_all(function(x){ifelse(is.nan(x), NA, x)}) %>% 
  ungroup() %>% 
  mutate(n.pos = n - n.miss) %>% 
  mutate(pos = ifelse(n.pos > 0, "pos", "neg")) %>% 
  mutate(percent_pos = (n.pos/n)*100)
wbe_tr
```

#### Percent positive technical reps for N1 only

```{r}
wbe_tr_n1 <- wbe_year2 %>% group_by(date, facility, target, biological_replicate) %>% 
  filter(target == "N1") %>% 
  dplyr::summarise(
    n=n(), 
    n.miss=sum(is.na(ct)), 
    ct.mean=mean(ct,na.rm=T), 
    ct.sd=sd(ct,na.rm=T)
  ) %>% 
  mutate_all(function(x){ifelse(is.nan(x), NA, x)}) %>% 
  ungroup() %>% 
  mutate(n.pos = n - n.miss) %>% 
  mutate(pos = ifelse(n.pos > 0, "pos", "neg")) %>% 
  mutate(percent_pos = (n.pos/n)*100)
wbe_tr_n1
```

#### Percent positive N2 only

```{r}
wbe_tr_n2 <- wbe_year2 %>% group_by(date, facility, target, biological_replicate) %>% 
  filter(target == "N2") %>% 
  dplyr::summarise(
    n=n(), 
    n.miss=sum(is.na(ct)), 
    ct.mean=mean(ct,na.rm=T), 
    ct.sd=sd(ct,na.rm=T)
  ) %>% 
  mutate_all(function(x){ifelse(is.nan(x), NA, x)}) %>% 
  ungroup() %>% 
  mutate(n.pos = n - n.miss) %>% 
  mutate(pos = ifelse(n.pos > 0, "pos", "neg")) %>% 
  mutate(percent_pos = (n.pos/n)*100)
wbe_tr_n2
```

### Calculate percent positive per biological replicate

`pos_rate_tr` will calculate the percent of positive technical replicates, averaged across biological replicates, within each sample's targets (N1 and N2)

`percent_pos_bio` will calculate percent of positive biological replicates for each target

```{r}
wbe_br <- wbe_tr %>%
  group_by(date, facility, target) %>% 
  dplyr::summarise(
    n.bio = n(),
    n.bio.non.miss = sum(!is.na(ct.mean)),
    n.bio.miss = sum(is.na(ct.mean)),
    n.total = sum(n), 
    n.total.miss = sum(n.miss), 
    bio.ct.mean = mean(ct.mean, na.rm = T), 
    bio.ct.sd = sd(ct.mean, na.rm=T), 
    pos_rate_tr = mean(percent_pos)) %>%
  mutate(percent_pos_bio = (n.bio.non.miss/n.bio)*100) %>% 
  mutate_all(function(x){ifelse(is.nan(x), NA, x)}) %>% 
  ungroup()
wbe_br
saveRDS(wbe_br, here("data/processed_data/prop_pos_bio.rds"))
```

### Calculate percent positive per sample

`pos_rate_tr` calculates the percent of positive technical replicates, averaged across all biological replicates and averaged across both targets

`prop.pos` calculates percent of positive targets for each sample

```{r}
wbe_samp <- wbe_br %>%
  group_by(date, facility) %>% 
  dplyr::summarise(
    n.pos = sum(!is.na(bio.ct.mean)),
    n.miss = sum(is.na(bio.ct.mean)),
    pos_rate_tr = mean(pos_rate_tr)) %>%
   mutate(prop.pos = n.pos/(n.pos+n.miss)*100) %>% 
  #mutate_all(function(x){ifelse(is.nan(x), NA, x)}) %>% 
  ungroup()

wbe_samp
saveRDS(wbe_samp, here("data/processed_data/prop_pos_sample.rds"))
```

### Calculate percent positive per collection date

```{r}
wbe_coll <- wbe_samp %>%
  group_by(date) %>% 
  dplyr::summarise(
    n.pos = sum(!is.na(n.pos)),
    n.miss = sum(is.na(n.miss)),
    pos_rate_tr = mean(pos_rate_tr)) %>%
   mutate(prop.pos = n.pos/(n.pos+n.miss)*100) %>% 
  #mutate_all(function(x){ifelse(is.nan(x), NA, x)}) %>% 
  ungroup()
```

### Combine all data: wastewater, clinical cases, percent positive

```{r}
final_data <- full_join(all_data, wbe_samp, by = c("date", "facility")) %>% 
  mutate(log10_VL = log10(total_copies), percent_pos_test = prop_pos*100) %>% 
  drop_na(facility)

saveRDS(final_data, here("data/processed_data/final_combined_data.RDS"))

```

### Summarize data across treatment facilities then average across targets

Group by: Date, Collection number

Keep: total copies, pos_rate_tr, log10 total copies, % positive tests

```{r}
final_data_sum <- final_data %>% group_by(date, collection_num, target) %>% 
  summarize(sum_total_copies = sum(total_copies),
            pos_rate_tr = mean(pos_rate_tr),
            percent_pos_test = mean(percent_pos_test),
            case_7dma = mean(pcr_pos_7dma),
            pcr_pos = mean(pcr_pos),
            test_admin = mean(pcr_tests)) %>% 
  ungroup(target) %>% 
  summarize(avg_VL = mean(sum_total_copies),
            pos_rate_tr = mean(pos_rate_tr),
            percent_pos_test = mean(percent_pos_test),
            case_7dma = mean(case_7dma),
            pcr_pos = mean(pcr_pos),
            test_admin = mean(test_admin)) %>% 
  ungroup() %>% 
  mutate(log10_VL = log10(avg_VL)) %>% 
  drop_na()

table1::label(final_data_sum$log10_VL) <- "Log10 Viral Load"
table1::label(final_data_sum$pos_rate_tr) <- "Assay Positivity"
table1::label(final_data_sum$case_7dma) <- "Clinical Case Seven-Day Moving Average"
table1::label(final_data_sum$pcr_pos) <- "Daily Reported Clinical Cases"
table1::label(final_data_sum$percent_pos_test) <- "PCR Test Positivity"

saveRDS(final_data_sum, here("data/processed_data/final_sum_combined_data.RDS"))

final_data_sum

```

```{r}

table1 <- (table1(~ log10_VL + pos_rate_tr + case_7dma + pcr_pos + percent_pos_test, data=final_data_sum)) 
saveRDS(table1, here("results/manuscript_table1.rds"))
```