### 2023-04-25 code/modeling_code/dont_run/2_VL_7DMA_model.Rmd

---
title: "VL_7DMA_model"
author: "Leah Lariscy"
date: "2023-03-17"
output: html_document
---

## Predictor and Outcome

Predictor: Log10 wastewater viral load

Outcome: Clinical case seven day moving average

## Load packages and data

```{r}
library(here)
library(tidyverse)
library(rsample) #Data split
library(tidymodels)
library(rpart) #Model fit
library(ranger) #Model fit
library(glmnet) #Model fit
library(rpart.plot)  #viz of decision tree
library(vip) #viz of variable importance plots
library(ggpmisc) #for adding linear regression to plots
```

```{r}
train_data <- readRDS(here("data/modeling_data/training_data.RDS")) %>% 
  as.data.frame()

test_data <- readRDS(here("data/modeling_data/testing_data.RDS")) %>% 
  as.data.frame()
```

## Define the linear regression model

The outcome is continuous, so we will use a linear regression model

```{r}
lm_mod <- linear_reg() %>% 
  set_engine("lm") %>% 
  set_mode("regression") %>% 
  translate()
```

## Define the null model

```{r}
null_mod <- null_model() %>% 
  set_engine("parsnip") %>% 
  set_mode("regression") %>%
  translate()
```

## Create recipe using training data

This is where we will set VL as predictor and 7DMA as outcome

```{r}
recipe_7dma <- recipe(case_7dma ~ log10_VL, data = train_data) %>% 
  step_nzv(all_nominal()) %>% 
  step_dummy(all_nominal())
```

## Create null model workflow

```{r}
workflow_null <- workflow() %>% 
  add_model(null_mod) %>% 
  add_recipe(recipe_7dma)
```

## Create training data workflow

```{r}
workflow_7dma <- workflow() %>% 
  add_model(lm_mod) %>% 
  add_recipe(recipe_7dma)
```

## Create resampling, 5x5

```{r}
folds <- vfold_cv(train_data, v = 5, repeats = 5, strata = case_7dma)
```

## Fit model to training data and null model resampling

```{r}
set.seed(626)
fit_7dma_cv <- fit_resamples(workflow_7dma, resamples = folds)
fit_null_cv <- fit_resamples(workflow_null, resamples = folds)

collect_metrics(fit_7dma_cv)
collect_metrics(fit_null_cv)
```

## Fit model to training data

```{r}
fit(workflow_7dma, data = train_data)
```

## Use model on the training data

```{r}
aug_train <- augment(fit_7dma_cv, train_data)
aug_train %>% select(case_7dma, .pred)
```

## Use RMSE and RSQ to evaluate model performance

```{r}
rmse <- aug_test %>% rmse(truth = case_7dma, .pred)
rsq <- aug_test %>% rsq(truth = case_7dma, .pred)
VL_7DMA_metrics <- full_join(rmse, rsq)

saveRDS(VL_7DMA_metrics, here("data/modeling_data/metrics/VL_7DMA_performance.RDS"))

VL_7DMA_metrics

```

## Plot actual value vs predicted value

```{r}
aug_test %>% ggplot(aes(case_7dma, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")
ggsave(here("data/modeling_data/metrics/VL_7DMA_predict.png"))
```

## Plot residual values

```{r}
aug_residual <- aug_test %>% mutate(residual = case_7dma - .pred)
aug_residual %>% ggplot(aes(.pred, residual)) +
  geom_point()

ggsave(here("data/modeling_data/metrics/VL_7DMA_residual.png"))
```

```{r}
aug_test %>% ggplot(aes(date)) +
  geom_line(aes(y=.pred)) +
  geom_line(aes(y=case_7dma), color = 'red')
```

## Use model on testing data

```{r}
aug_test2 <- augment(fit_7dma, test_data)
aug_test2 %>% select(case_7dma, .pred)

rmse <- aug_test2 %>% rmse(truth = case_7dma, .pred)
rsq <- aug_test2 %>% rsq(truth = case_7dma, .pred)
huber <- aug_test2 %>% huber_loss(truth = case_7dma, .pred)
VL_7DMA_test_metrics <- bind_rows(rmse, rsq, huber)

VL_7DMA_test_metrics
```

## Use model on testing dates

```{r}
aug_test3 <- augment(fit_7dma, test_dates)
aug_test3 %>% select(case_7dma, .pred)

rmse <- aug_test3 %>% rmse(truth = case_7dma, .pred)
rsq <- aug_test3 %>% rsq(truth = case_7dma, .pred)
huber <- aug_test3 %>% huber_loss(truth = case_7dma, .pred)
VL_7DMA_test_dates_metrics <- bind_rows(rmse, rsq, huber)

VL_7DMA_test_dates_metrics
```

```{r}
aug_test3 %>% ggplot(aes(x=date)) +
   geom_line(aes(y=case_7dma), color = "#1c246c", size = 2) +
   geom_line(aes(y=.pred), color = "#ec4998", size = 2) +
   theme_light() +
     ylab(label = "Predicted 7-Day Moving Average") +
  xlab(label = "Date") +
  theme_light()
```