### 2023-04-26 code/modeling_code/dont_run/PPBoth_DailyPosTest_model.Rmd

---
title: "PP_7DMA_model"
author: "Leah Lariscy"
date: "2023-03-17"
output: html_document
---

## Predictor and Outcome

Predictor: Percent positive technical replicates per sample

Outcome: Clinical case seven day moving average

## Load packages and data

```{r}
library(tidyverse)
library(tidymodels)
library(here)
```

```{r}
train_data <- readRDS(here("data/modeling_data/training_data_both.RDS"))
test_data <- readRDS(here("data/modeling_data/testing_data_both.RDS"))
```

## Define the model

The outcome is continuous, so we will use a linear regression model

```{r}
lm_mod <- linear_reg()
```

## Create recipe using training data

This is where we will set PP as predictor and 7DMA as outcome

```{r}
recipe_7dma <- recipe(pcr_pos ~ percent_pos, data = train_data)
```

## Create workflow

```{r}
workflow_7dma <- workflow() %>% 
  add_model(lm_mod) %>% 
  add_recipe(recipe_7dma)
```

## Create model using workflow

```{r}
set.seed(828)
fit_7dma <- workflow_7dma %>% 
  fit(data = train_data)

tidy(fit_7dma)
```

## Use model on the training data

```{r}
aug_test <- augment(fit_7dma, train_data)
aug_test %>% select(pcr_pos, .pred)
```

## Use RMSE and RSQ to evaluate model performance

```{r}
rmse <- aug_test %>% rmse(truth = pcr_pos, .pred)
rsq <- aug_test %>% rsq(truth = pcr_pos, .pred)
huber <- aug_test %>% huber_loss(truth = pcr_pos, .pred)
PP_7DMA_metrics <- bind_rows(rmse, rsq, huber)

saveRDS(PP_7DMA_metrics, here("data/modeling_data/metrics/PP_7DMA_performance.RDS"))

PP_7DMA_metrics

```

## Plot actual value vs predicted value

```{r, message=FALSE}
aug_test %>% ggplot(aes(pcr_pos, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")
ggsave(here("data/modeling_data/metrics/PP_7DMA_predict.png"))
```

## Plot residual values

```{r, message=FALSE}
aug_residual <- aug_test %>% mutate(residual = pcr_pos - .pred)
aug_residual %>% ggplot(aes(.pred, residual)) +
  geom_point()

ggsave(here("data/modeling_data/metrics/PP_7DMA_residual.png"))
```

## Plot 7DMA and prediction temporally

```{r}
aug_test %>% ggplot(aes(x=date)) +
  geom_point(aes(y=pcr_pos), color = "blue") +
  geom_line(aes(y=pcr_pos), color = "blue") +
  geom_point(aes(y=.pred), color = "green") +
  geom_line(aes(y=.pred), color = "green")

```

## Use model on testing data

```{r}
aug_test2 <- augment(fit_7dma, test_data)
aug_test2 %>% select(pcr_pos, .pred)
```

```{r}
aug_test2 %>% ggplot(aes(x=date)) +
  geom_point(aes(y=pcr_pos), color = "blue") +
  geom_line(aes(y=pcr_pos), color = "blue") +
 geom_point(aes(y=.pred), color = "green") +
  geom_line(aes(y=.pred), color = "green")
```

## Evaluate model performance on testing set

```{r}
rmse <- aug_test2 %>% rmse(truth = pcr_pos, .pred)
rsq <- aug_test2 %>% rsq(truth = pcr_pos, .pred)
huber <- aug_test2 %>% huber_loss(truth = pcr_pos, .pred)
PP_7DMA_test_metrics <- bind_rows(rmse, rsq, huber)

saveRDS(PP_7DMA_test_metrics, here("data/modeling_data/metrics/PP_7DMA_test_performance.RDS"))

PP_7DMA_test_metrics
```

## Use model on testing dates

```{r}
aug_test3 <- augment(fit_7dma, test_dates_both)
aug_test3 %>% select(pcr_pos, .pred)
```

## Evaluate model performance on testing dates

```{r}
rmse <- aug_test3 %>% rmse(truth = pcr_pos, .pred)
rsq <- aug_test3 %>% rsq(truth = pcr_pos, .pred)
huber <- aug_test3 %>% huber_loss(truth = pcr_pos, .pred)
PP_7DMA_BA5_metrics <- bind_rows(rmse, rsq, huber)

saveRDS(PP_7DMA_BA5_metrics, here("data/modeling_data/metrics/PP_7DMA_BA5_performance.RDS"))

PP_7DMA_BA5_metrics
```

```{r}
aug_test3 %>% ggplot(aes(x=date)) +
  geom_line(aes(y=pcr_pos), color = "#621940", size = 2) +
  theme_light()

```

```{r, message=FALSE}
 aug_test3 %>% ggplot(aes(x=date)) +
   geom_line(aes(y=pcr_pos), color = "#1c246c", size = 2) +
   geom_line(aes(y=.pred), color = "#ec4998", size = 2) +
   theme_light() +
     ylab(label = "Predicted Daily Clinical Cases") +
  xlab(label = "Date") +
  theme_light()

ggsave(here("figures/modeling/predicted_DailyCase.png"))
```